Lang-gen model simulations

### 2018-02-05 analysis/model-simulations.Rmd
simply fig1, modularize code to analysis script for model simulation, cache model sims

---
title             : "The language of generalization: Probability, vagueness, and interaction"
shorttitle        : "The language of generalization"

author: 
  - name          : "Michael Henry Tessler"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Bldg. 420, Rm. 316, Stanford, CA 94305"
    email         : "mtessler@stanford.edu"
  - name          : "Noah D. Goodman"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

header-includes:
  - \usepackage{tabularx}
  - \usepackage{multicol}
  - \usepackage{wrapfig}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{graphicx}
  - \usepackage{subcaption}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}

author_note: >
  This manuscript is currently in prep. Comments or suggestions should be directed to MHT.

abstract: > 
    Generalizations are central to human understanding and language provides simple ways to convey them (e.g., "Birds fly").
    The language of generalization is ubiquitous in everyday discourse and child-directed speech. 
    Yet the meaning of these expressions is logically puzzling (e.g., not all birds fly) and has resisted precise formalization.
    The major issue in formalizing generalization in language is determining which statements are true or which are false.
    Using an information-theoretic probabilistic model, we explore the hypothesis that the meaning of these linguistic expressions is *simple but underspecified*, and that general communicative principles can be used to establish a more precise meaning in context. 
    To test this theory, we examine endorsements of generalizations about three different domains: categories (*generic language* e.g., "Birds fly"), events (*habitual language* e.g., "John runs"), and causes (*causal language* e.g., "Staring at the sun makes you go blind"). 
    Across these diverse domains, we find that our model explains the wide variance in human endorsements of language conveying generalization, while simpler models fall short.
    Ours is the first formal theory that makes precise, quantitative predictions about human understanding of the language of generalization.
    The results demonstrate that the context-sensitivity of such language emerges from the interaction of an underspecified meaning with diverse beliefs about properties.

keywords          : "genericity, generalization, generics, pragmatics, semantics, bayesian modeling"
wordcount         : "20000"

bibliography      : ["generics.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

\newcommand{\denote}[1]{\mbox{ $[\![ #1 ]\!]$}}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\definecolor{Red}{RGB}{255,0,0}
\definecolor{Green}{RGB}{10,200,100}
\definecolor{Blue}{RGB}{10,100,200}

\newcommand{\mht}[1]{{\textcolor{Blue}{[mht: #1]}}}
\newcommand{\ndg}[1]{{\textcolor{Green}{[ndg: #1]}}}
\newcommand{\red}[1]{{\textcolor{Red}{#1}}}

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = F, fig.path='figs/',
                      echo=FALSE, warning=FALSE, cache=T, message=FALSE, sanitize = T)

@@

<!-- 
possible titles:
- Communicating generalizations
- The language of generalization
- The language of generalization: Probability, vagueness, and interaction
-->

# Introduction

Learning that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to thrive in our open-ended world.
Yet such knowledge can be difficult to acquire: Relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
It thus is important that language allow us to communicate such *generalizations* to each other.
By sharing generalizations, we flourish collectively without individually needing to taste potentially-poisonous plants or personally witness many lightning strikes.
Flexibly communicating generalizations from one generation to the next may be a foundational mechanism behind the faithful transmission of knowledge necessary for cumulative cultural evolution [@Tomasello1999; @Henrich2015].

<!-- Generalizations can be acquired from the world by drawing inductive inferences from observations [@Baldwin1993; @Xu2008; @Dewar2010]. \ndg{cite older stuff. eg hume} -->
<!-- Cooperative teachers can facilitate inductive learning by supplying helpful observations to the learner [@Csibra2009; @Tomasello1999; @Butler2012]. -->
<!-- At around a child's first birthday, however, the learner gains access to a new, human-unique source of data: language. -->
<!-- Minimally, language fosters generalizations by bringing attention to objects being of the same kind through the use of labeling [e.g., "This is a dog"; @Gelman1986; @Gopnik2000; @Graham2004; @Markman1989; @Schulz2008]. -->
<!-- Objects have many properties though; when speakers attach properties to kind-labels (e.g., "Dogs bark"), listeners hear a generalization. -->
<!-- \ndg{this par doesn't really do much} -->

The language of generalization is often studied via *generic language* (or, *generics*): simple statements predicating properties of categories [e.g., "Dogs are friendly"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008].
Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
In contrast to *particular statements* (e.g., "Rufus is friendly") which describe individual entities or situations and can be verified through observation, generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dogs}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
Generics are just one case in the language of generalization, however; events can be described in generalization (*habitual language*; e.g., "Mary swims after work.") as well as causal relationships (*causal language*; e.g., "Staring at the sun makes you go blind.").
Every language, it is believed, can express generalization [@Behrens2005; @Carlson1995] and is central to the growth of conceptual knowledge [@Gelman2004].
Considering generalization as a singular phenomenon in language, thus, promises to provide understanding of a wide variety of linguistic expressions with powerful implications.

<!-- Generic statements ascribing properties to categories are just one kind of generalization found in language.  -->
<!-- Individual entities can also be talked about in generalization, using so-called  -->
<!-- <!-- Recent evidence suggests talking about the other in generalization (using so-called *generic You*) can help make meaning from difficult situations   [e.g., "You never know what will happen on a blind date."; @Orvell2017]. -->
<!-- Communicating about causality can also take the form of communicating a generalization about a causal event  -->
<!-- What these statements have in common is that they convey generalizations, or exhibit *genericity* [@Carlson1977; @Carlson1995]. -->

Despite their ubiquity and relative simplicity, there have been few formal models of the language of generalization and none that have made precise, quantitative predictions to match human intuitions. 
The major challenge in formalization is determining *what makes the statements true or false*.
When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark".
But the property does not apply to all dogs: Basenji dogs are barkless^[
http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html
], for example.
The generalization could signal something weaker like "Most dogs bark".
However, most mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only quantifier that would permit such a statement would be "some" (e.g., "Some mosquitos carry malaria").
But weakening the generalization to mean "some" is too permissive: some robins are female (in fact, half of them are), but asserting "Robins are female" seems odd.
Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say.

These observations have led to the conclusion that the prevalence of the feature (or the probability of a member of the category having the property e.g.,  $P(x \text{ barks} \mid x \text{ is a dog})$) is not the only parameter in the meaning of generalizations.
Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012]. 
"Bishops move diagonally" not because bishops tend to move diagonally but rather because those are the rules of the game.
It's been suggested that the mind has cognitive primitives that directly dictate what generalizations are true or false.
One such primitive is the ability to priviledge *characteristic features*: For an animal, their mode of reproduction is a characteristic feature. 
Thus, whatever their mode of reproduction may be will make a true generic statement (e.g., "Robins lay eggs"), regardless of the actual prevalence of the feature.
*Horrifying* or *striking* information is also priviledged and creates true generics, which is why "Mosquitos carry malaria" is true

Yet, generics convey more than just rich, conceptual relations [@Nickel2008; @Nickel2016; @Sterken2015]. 
Purely arbitrary relationships can result in true generics.
<!-- Arbitrary relationships are also expressed with generics:  -->
"Barns are red" because barn-owners tend to paint their barns red; were barn-owners to decide tomorrow to switch to blue, then "Barns are red" would be false.
Ravens and toasters have no priviledged relationship to each other, but the statement "Ravens are bigger than toasters" is still a true generalization.
Conceptual accounts must be supplemented with the proviso that, in some situations, the actual prevalence of the feature matters for making generalizations true.

What we are left with is hybrid theories, which try to classify different kinds of generalizations and the various inferences they support [e.g., @Prasada2013].
But insofar as there is a single class of linguistic expressions that convey generalization, there should be something common to them all: a general, unified semantic core.
In this paper, we propose such a semantic core for the language of generalization.
To do so, we look to probability, the universal currency of belief and a useful representation for human generalization from observations [@Shepard1987; @Tenenbaum2001].
Given that certain probabilities are the result of inductive generalization, it would make sense that these are the object of communication in the language of generalization. 
To formalize a model of endorsement (i.e., what makes the statements true or false), we consider the information-theoretic function of a generalization by explicitly modeling the background knowledge a language interpreter would use when understanding such language.
We create an endorsement model based on this interpretation hypothesis, providing the first formal theory that makes quantitative predictions about what makes generalizations in language true or false.
The assumptions of the theory, and in turn, the conditions under which the theory may be meaningfully tested are made explicit by its instantiation in a formal model.
Though concepts such as *striking properties* are not explicitly formalized in our model, our general cognitive modeling framework provides natural avenues for future integration.

<!-- The test of a unified semantic core is its ability to explain what makes the statements true or false. -->

<!-- It is thus argued that there are qualitatively different kinds of generic statements (i.e., "striking", "characteristic", "statistical"), each with their own conditions on what makes them true. -->
<!-- This diversity is interesting, but what is common across generics? -->
<!-- Is there some core meaning to *genericity*, a general unified principle that explains what makes generic statements true or false? -->
<!-- Finally, formalizing such a theory is crucial to make explicit the relevant assumptions of the psychological or linguistic hypothesis meant to match human intuitions; this, in turn, makes the theory more testable.  -->
<!-- We formalize such a semantics in a general cognitive modeling framework, and thus incorporate generalizations in language into a larger class of models designed to formalize concepts and common-sense reasoning [@Goodmanconcepts]. -->

<!-- Specifically, we formalize an account where the core meaning of a generalization in language is a *simple but underspecfied* function of probability and where context helps resolve the underspecification. -->
<!-- In its simplest form, context takes the shape of interlocutors' shared beliefs about the relevant properties or categories under discussion [cf., @Grice1975; @Clark1996; @Levinson2000]. -->
<!-- We then ask what a rational speaker---who can only either endorse the generalization or stay silent---would say in different contexts; this is our formal model of endorsement. -->
<!-- We show how the decision to endorse depends in systematic ways on the shared prior knowledge of the interlocutors. \mht{"interlocutors" may be confusing here...} -->
<!-- We test this theory in three distinct case studies---generalizations about categories, events, and causes---by both measuring and manipulating prior knowledge, showing that it is strongly and causally related to the endorsement of generalizations in the ways predicted by our model. -->

<!-- We show that in each domain, our general theoretical framework explains endorsements about generalizations conveyed in language, while simpler models fall short.  -->

<!-- We think prevalence may have been cast aside too quickly for two main reasons: (i) in previous empirical work, prevalence has been rather narrowly defined, as frequency, where in fact it can be more generally seen as probability (ii) the backround knowledge that adult speakers, whose intuitions are consulted to establish a theory of meaning, may bring to the table has not been explicitly considered or formalized in studies and theoretical work into the language of generalization. -->

<!-- We propose a formal, unified account of generalization in language that can make precise, quantitative predictions about these simple and ubiquituous utterances. -->
<!-- We do so by first considering a much more well-studied form of generalization: generalization from observations.  -->
<!-- The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is has proven to be a good model of human generalization . -->

<!-- We formalize this using a rational model of communication --- the Rational Speech Act (RSA) theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016]. -->
<!-- Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals.  -->
<!-- Heretofore, however, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013]. -->
<!-- The language of generalization does not make reference to entities in the world or concrete situations, but rather refer to unobservable concepts in the head. -->

<!-- We relate known measurements on factors regarding generic language to each other via our model. -->
<!-- We then go on to manipulate the relevant factors, showing they are causally related via our model to generic endorsement and interpretation.  -->

The paper is organized as follows. 
In Section 1, we formalize the hypothesis that the core of meaning of a generalization in language is simple but *underspecified*, articulate a thory of relevant background knowledge used to understand a generalization in language, and develop a computational model of endorsement around these components.
The endorsement model's predictions depend upon the background knowledge and the probability communicated, which we demonstrate via simulation.
Sections 2 - 4 present our three empirical case studies: generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
Across these three case studies, we both measure and manipulate background knowledge and the probability communicated, measuring their effects on endorsement. 
We find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements, where alterantive models fall short.
This work opens the door to understanding generalizations in pragmatic language understanding contexts and building models of learning from language.

<!-- (cumulative $r^2(N) = 0.XX$). -->
<!-- In the final section of the paper, we discuss the generality of our approach and pave a path to unifying the extant psychological, linguistic, and philosophical literatures on generalization in language. -->
<!-- We also show how our endorsement model is a special case of a speaker model in a rational framework of communication.  -->

# Computational Framework

Generalizations are used to make predictions about properties of instances that an agent has yet to experience [@HumeTHN].
People readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find certain people typing.
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$.
This prediction can be described by a conditional probability: $p_{fk} = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ (formally, be in the set of things that have $f$) given that it is in $k$, which we will refer to as the **prevalence**.
The targets of our predictions can vary widely: They may be objects (a dog), events (a coffee drink), or more fine grained types (a person on a given day).
The properties also vary (e.g., having fur, causing jitters, typing). 
Yet the mathematical description of the inductive inference is always given by the probability $p_{fk}$.

How does one acquire such an inductive belief?
Observing instances of the category $k$ and noting how many of them have $f$ can support prediction to new instances when integrated with one's prior expectations about the property [cf., @Gelman2003; @Keil1992].
For example, if you observe creatures that \textsc{have two legs}, you might infer that all creatures of that category have two legs.
Observing new creatures with \textsc{broken wings}, however, leads to a much weaker prediction about other instances of the category having the property [@Nisbett1983].
This kind of domain-specific knowledge about properties can be represented formally as a probability distribution over the prevalence $p_{f}$: $P(p_{f})$ [@Kemp2008], and observational data $d$ can update one's beliefs via Bayes' Rule: $P(p_{f} \mid d) \propto P(d \mid p_{f}) \cdot P(p_{f})$

Learning from observational evidence can only take you so far, though. 
To learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*), one would much rather learn from others.
Fortunately, language has simple ways of encoding generalizations.

<!-- Here, $P(d \mid h)$ is the observer's generative model of the observed data.  -->
<!-- It is conditional on $h$ because the probability of various direct observations depends upon the prevalence that the observer assumes.  -->
<!-- Generalization from observational evidence is limited, however, in its capacity to describe how the mind builds complex theories of the world or how culture accumulates across generations. -->
<!-- How do we acquire these generalizations? -->
<!-- For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations. -->
<!-- The human capacity to generalize from observations, deriving $h$ from example $x$s, has been studied extensively in cognitive and developmental psychology [@Nisbett1983; @Baldwin1993; @Heit2000]. -->
<!-- From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007]. -->
<!-- Such inductive generalization can be described by the Bayesian probability calculus: An observer's initial beliefs or state of knowledge---called a prior belief distribution---is updated through experience (i.e., after making observations) and becomes a posterior belief distribution (what they belief *a posteriori*).  -->
<!-- This posterior distribution becomes the observer's prior for the next observation she makes: it thus informs her predictions about unobserved instances. -->

<!-- As a concrete example, consider a learner investigating the feature \textsc{barks} ($f$) for the category \textsc{dog} ($k$). -->
<!-- $h$ describes the learner's subjective probability that the next dog she encounters will bark^[At the moment, we assume the feature \textsc{barks} is an all-or-none Boolean property. Later, we will see how the presence of this feature itself can be thought of as a generalization over events of barking.], or more abstractly, her belief about the barking capacity of dogs. -->
<!-- That is, $h = P(x \text{ barks} \mid x \text{ is a dog})$. -->
<!-- The learner does not know $h$, but has some idea about what $h$ could be; this is represented by $P(h)$, the learner's prior beliefs about how likely it is that $x$ will bark given that $x$ is a dog. -->

<!-- The observer goes around the world, encountering dogs and taking note of how many bark. -->
<!-- This could be represented as a number of positive examples $d$ out of a number of total examples $n$.  -->
<!-- With the weak assumption that the dogs she has observed were all independent and identically-distributed (*i.i.d.*) random samples from the category of dogs, the likelihood of her observations are given by the Binomial distribution.  -->
<!-- That is, $d \sim  \text{Binomial}(n, h)$, where $d$ is the number of dogs she has seen that bark, $n$ is the total number of dogs she has seen, and $h$ is the prevalence of barking among dogs. -->
<!-- This is called a likelihood function: It determines how well a given value of $h$ predicts the observed $d$.  -->
<!-- Given a formal description of the learner's state of knowledge before any observations $P(h)$ and the learner's assumptions about how the data she observed were generated (e.g., the *i.i.d.* assumption), what a learner should believe after having seen the data is proportional to the product of these two terms, as given by Bayes' Rule: -->

<!-- $P(h \mid d)$ is then the strength of the learner's belief in the next dog she encounters barking. -->
<!-- This simple idea of integrating observational evidence with prior knowledge is surprisingly powerful in explaining a multitude of diverse phenomena, from perception to word learning to inductive reasoning [e.g., @Tenenbaum2006; @Tenenbaum2011]. -->
<!-- \ndg{what is this par doing for us? delete?} -->
<!-- Generalization from examples shows complex sensitivites to the kind of observations, and even how those observations came to be observed in the first place. -->
<!-- For instance, observing that an indigenous person living on a remote island *is brown and obese* provides considerably more information about the skin color of other islanders than about their weight [@Nisbett1983]. -->
<!-- This subtlety can be modeled by incorporating heiarchical structure into the prior belief distribution of a Bayesian model, allowing the model to represent and reason flexibly about different kinds of properties [@Kemp2008]. -->
<!-- Observations that are generated by an intentional agent can lead to similarly subtle inferences, like deconfounding otherwise ambiguous causal evidence [@Goodman2009]. -->
<!-- Yet generalizations from observational evidence is limited in its capacity to describe how the mind builds complex theories of the world or how culture accumulates across generations. -->
<!-- Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *staring at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*). -->
<!-- How do we acquire these generalizations? -->
<!-- For abstract, generalizable knowledge to remain in a culture, it must be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations. -->

## Generalization from language

The language of generalization manifests when a property is predicated of a category: "Ks F" (e.g., "Dogs bark", "Staring at the sun makes you go blind"). 
Generalization from observations can be described as a predictive probability; it is intuitive then to posit that same quantity will be at the core of a semantic theory of the language of generalization.
To connect to the semantic tools of *truth values* [@Montague1973], we must relate the scalar quantity of probability to a Boolean value.
One simple way is with a *threshold semantics*: An utterance is true if the relevant quantity is above a threshold.
A threshold semantics for a generalization about category K with property F would then apply to the *prevalence* $p_{fk}$:
$\denote{K \text{ } F}(p_{fk}, \theta) = \{p_{fk} >\theta\}$

Various attempts of this kind have been made before [e.g., @Cohen1999].
We distinguish our proposal from previous ones by emphasizing the cognitive underpinnings of $p_{fk}$, as well as the formalization into a model that makes precise, quantitative predictions about human endorsements.
$p_{fk}$ describes a latent belief and should not be confused with *frequency*, which can inform a belief but is not itself the belief.

<!-- We emphasize that this account is based on a probability $h$. -->
<!-- The literature on generic language has often focused on *prevalence*---either the actual or believed proportion of members of a category with the property---which we distinguish from the prevalence $h$.^[ -->
<!--   There has also been substantial discussion of *cue validity* , which we come back to below. -->
<!-- ] -->
<!-- Prevalence, an ostensibly observable quantity, can inform a person's beliefs (e.g., via Bayesian belief updating) but it is not the belief itself.  -->
<!-- Probability is a measure of belief.  -->

### Literal interpretation model

Quantifier statements can be described using a similar threshold semantics: $\denote{some \text{ } K \text{ } F}(p_{fk}) = \{p_{fk} > 0\}; \denote{all \text{ } K \text{ } F} = \{p_{fk} = 1\}$.
So what threshold $\theta$? should be used for the generalizations? 
Unlike quantified language which have fixed-thresholds, we posit that the language of generalizations is *underspecfied* with respect to the actual truth-functional threshold $\theta$.
That is, we hypothesize that $\theta$ is in part determined by contextual, in a way analagous to how gradable adjectives like "tall" have contextually-determined thresholds [e.g., what counts as "tall" for a four-year-old is different than what counts as "tall" for an adult; @Kennedy2007; @Lassiter2013].

We formalize this underspecified semantics in a probabilistic model, where the underspecifed threshold $\theta$ is sampled from a uniform prior distribution over possible thresholds $P(\theta)$ [cf., @Lassiter2013].
After making explicit the relevant background knowledge about the property $P(p_f)$, the model becomes:

\begin{eqnarray}
Lit(p_{fk}, \theta \mid u) &\propto& {\delta_{\denote{u}(p_{f}, \theta)} \cdot P(p_{f}) \cdot P(\theta)} \label{eq:L0}
\end{eqnarray}

Equation \ref{eq:L0} ($Lit$) is a literal interpretation model, that updates its beliefs about the prevalence of a feature in the category described $p_{fk}$ according to the truth-functional meaning of the utterance $u_{KF}$. 
Formally, the truth-functional meaning is represented by the Kronecker delta function  $\delta_{\denote{u}(p_{fk}, \theta)}$ that returns probabilities proportional to $1$ when the utterance is true (i.e., when $p_{fk} > \theta$) and $0$ otherwise.

\begin{eqnarray}
\delta_{\denote{u_{KF}}(p_{fk}, \theta)} &\propto  & \begin{cases}
1 & \text{if } p_{fk} > \theta \\
0 & \text{otherwise}
\end{cases}\label{eq:delta}
\end{eqnarray}

### Endorsement model

Given the minimal model of literal interpretation described above, how do we decide when a statement conveying a generalization is true or not?
We consider the notion of felicity or endorsement from the information-theoretic perspective: Would a generalization convey sufficient information to warrant the cost of production? 
We formalize this by considering how well the generalization would convey the prevalence $p_{fk}$ to the $Lit$ model (Eq. \ref{eq:L0})^[
  A more general version of this model can relax the assumption that the endorsement model has access to a specific prevalence $p_{fk}$ that it wants to communicate.
  Rather, the endorser may have a distribution over $p_{f}$ for the category $k$, corresponding to the endorser's uncertain beliefs about the prevalence. 
  In this situation, we would define the endorsement model decision to be with respect to the expected value of the informativity, which integrates over the endorsement model's belief distribution:
  $S(u \mid k) \propto \exp{(\lambda \cdot {\mathbb E}_{p_{fk}\sim P_{k}} \ln{ \int_{\theta} Lit(p_{fk}, \theta \mid u)} \diff \theta )}$
  For the empirical case studies described below, these two versions of the model make almost identical predictions.
].

\begin{equation} 
S(u \mid p_{fk}) \propto \exp{(\lambda \cdot \ln{ \int_{\theta} Lit(p_{fk}, \theta \mid u)  \diff \theta } )}
\label{eq:S1}
\end{equation}

The endorsement model simulates what the interpretation model would believe upon hearing an utterance $u$): $Lit(p_{fk}, \theta \mid u)$.
$S$ is not interested in communicating $\theta$ (in fact, it does not know $\theta$), and so integrates over the values of $\theta$.
We assume the decision to endorse is an approximately rational one, with degree of rationality given by $\lambda$.
The proportionality in Eq. \ref{eq:S1} implies normalization over a set of alternative utterances. 
We interpret the endorsement task (e.g., "true" vs "false"; "agree" vs. "disagree") as supplying two alternatives: *endorse* or *not*, where *not* we take to be a silent utterance that conveys no information content [cf., @Franke2014cogsci; @Degen2014]: $\denote{null}(x, \theta) = \text{True}$^[
  This "null" alternative  to the generalization ("Ks F") can be realized in at least two other ways: the negation of the generalization (i.e., "It is not the case that Ks F") or the negative generalization (i.e., "Ks do not F").
  All results reported are similar for these two alternatives, and we use the alternative of the "silent" utterance for simplicity.
]

Our endorsement model (Eq. \ref{eq:S1}) provides a straight-forward mapping from a *referent-category prevalence* $p_{fk}$  (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) to an endorsement probability for the corresponding generalization (e.g., "Robins lay eggs"): $S(u = u_{KF} \mid p_{fk})$. 
The model assumes a threshold semantics based on probability (prevalence) but does not require specifying the threshold *a priori* (note that there is no $\theta$ in the left-hand side of Eq. \ref{eq:S1}).
This model can be seen as a special case of a Rational Speech Act model [@Frank2012; @Goodman2016], where there is only a single level of recursion and the space of possible utterances is maximally constrained.

### Prevalence priors and referent-prevalence

Our model makes context-sensitive predictions about endorsing generalizations based on two model components: The prevalence priors $P(p_f)$ in Eq. \ref{eq:L0} and the referent-prevalence $p_{fk}$ in Eqs. \ref{eq:S1} and \ref{eq:delta}.
Referent-prevalence is the probability that an individual $x$ of the referent-category (e.g., "Dogs") will have the feature being predicated of the category (e.g., "...barks").
We will return and refine this definition in Case Study 2, when considering generalizations about events.
For now, the referent-prevalence is simply a probability that determines the truth-conditional semantics. 
Our hypothesis does not provide a mapping from the referent-prevalence directly to endorsement judgments, however, because $\theta$ is underspecified.

The uncertain, truth-conditional threshold $\theta$ is resolved by the literal interpretation model $Lit$ (Eq. \ref{eq:L0}) by integrating prior beliefs about the feature in question---the *prevalence prior* $P(p_f)$---with the truth-functional semantics.
Again, $P(\theta)$ is set to be a uniform distribution across all contexts, and thus, the generalization could in principle be felicitious with potentially any referent-prevalence. 
However, interpretation model's *a priori* beliefs $P(p_f)$ must be considered before the endorsement model can say whether or not it would endorse the generalization.

Beliefs about properties result from internal conceptual models of the world. 
These internal models have domain-specific features but should all give to the domain-general currency of inductive reasoning: probability.
One way of constructing prior distributions over prevalence, without appealing directly to domain-specific conceptual knowledge, is by considering the prevalence of the feature among different categories.
For example, many different kinds of animals *do not bark* (more precisely: have close to 0 exemplars who bark); thus, the prevalence prior for the feature *barks* $P(p_{barks})$ should have substantial probability mass near 0, because what one would expect for a randomly-sampled category is that it has 0 or close to 0 exemplars that bark.
Furthermore, among the categories that do have exemplars which bark (e.g., *dogs*, *seals*, ...), the probability is quite high that an exemplar will bark; thus, the prevalence prior should have substantial mass at relatively high values, potentially with quite a bit of uncertainty; Figure\ \@ref(fig:simulations)A; first row).
Conceptual differences can manifest in these prevalence priors in the form of multi-modal distributions; for example, the prevalence prior for *lays eggs* should similiarly have substantial probability mass near 0 (since most animal categories do not have egg-layers) but should have a secondary component peaked around 0.5, because among the animals with egg-layers, only the female members of the category have the property (Figure\ \@ref(fig:simulations)A; bottom row).

By encoding knowledge about other categories, the prevalence prior distributions are deeply connected to the construct of *cue validity*, or the probability that an entity with a certain property $f$ is a member of kind $k$: $P(x \in k \mid x \in f)$ (e.g., one's predictions about whether or not the entity is a robin, upon learning that it lays eggs). 
*Cue validity* implicitly encodes knowledge of other categories and has been shown to be  highly correlated with endorsing *generic sentences*, or generalizations about categories [@Khemlani2012].
Cue validity is defined for a particular (category, property) pair (as is referent-prevalence), and relates to referent-prevalence by Bayes' Rule:
\begin{equation}
P( k \mid  f) = \frac{P( f \mid  k) \cdot P( k)}{\sum\limits_{k' \in K} P( f \mid k') \cdot P( k')} \label{eq:cuevalidity}
\end{equation}
Since *cue validity* implicitly encodes knowledge of other categories (by summing over $k'$) and the prevalence prior explicitly encodes knowledge of other categories, how are these two constructs related?
In the prevalence prior $P(p_f)$, a category is represented only by its corresponding prevalence (i.e., a sample from the prevalence prior can be thought of as a category $k$, with the only information available about the category being the prevalence of the feature in that category $p_{fk}$). 
Thus, $P(p_{fk}) = P(k_{f})$.
The denominator of Eq. \ref{eq:cuevalidity} (which encodes knowledge of other categories), then, is the expected value (i.e., the mean) of the prevalence prior distribution: $\mathbb{E}[P(p_{f})] = \sum_{k' \in K} p_{fk'} \cdot P(p_{fk'} )$.
For a more thorough derivation of this relationship, see Appendix A.
The fact that the prevalence prior distribution relates only via its mean value to *cue validity* reveals a potential shortcoming of cue validity: By reducing a distribution to a point estimate, the structure of the distribution is lost. 
This is particularly problematic for prevalence distributions that encode conceptual knowledge via multimodal distributions (e.g., Fig. \@ref(fig:simulations)A; bottom row); the mean of a multimodal distribution is not a good representation of that distribution. 
We return to this point empirically in Case Study 1.

<!-- These distributions can be thought of as distributions over categories, displaying only the prevalence for the property in question. -->

<!-- F(left; filled colors) shows six prior distributions over feature-probabilities, representing the relevant background knowledge about each property. -->
<!-- For example,   -->

### Model simulations

The endorsement model defined in the previous section explicitly predicts that the variability in endorsements for generalizations in language can be explained by two factors: (i) the prevalence communicated in the category under discussion $p_{fk}$ in Eq. \ref{eq:S1} (e.g., $P(x \text{ lays eggs} \mid x \text{ is a robin})$) and (ii) the background knowledge about prevalences $P(p_f)$ (Eq. \ref{eq:L0}).
We simulate how these two factors change both literal interpretation and endorsement using schematic parameter values based on intuition from classic examples in the literature on generic language.
We implemented these and all subsequent Bayesian models in the probabilistic programming language WebPPL [@dippl].
The following simulations were implemented using the R package RWebPPL in the RMarkdown document that generated this paper text and figures.
That document, as well as all other models, analysis, data and links to experiments used in this paper can be found at \url{https://mhtess.github.io}.

<!-- depends upon the probability that utterance vs. silence accurate conveys to a naive listener a. -->
<!-- This probability of successful communication is given by $L(h \mid u, \theta)$ (Eq. \ref{eq:S1}) and depends upon the listener's *a priori* beliefs about the prevalence in question: $P(h)$ (Eq. \ref{eq:L0}), as well as the referent-category prevalence $h$ (Eq. \ref{eq:S1}). -->
<!-- Predictions for endorsement depend upon how well the listener's posterior on $h$  upon hearing a generalization matches an $h$ that the speaker wants to convey in comparison to how well the listener's prior matches (which would result from the speaker staying silent). -->
<!-- For purposes of illustration, we present different hypothetical priors on feature-probabilities $P(h)$, together with listener $L$ posterior distributions on $h$ after hearing utterances corresponding to (i) the underspecified threshold semantics as well as (ii) a comparison model where the threshold is fixed near-0, corresponding intuitively to the quantifier *some* (Figure \ref{fig:simulations}). -->

@@

```{r rsaInterpretationModels, eval = T, cache = T}
no.utterance.model <- '
var noUtteranceInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    return { state }
 }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
'

@@

```{r cache = F}
fig.sims.priors <- ggplot(sims.priors, aes(x = state, fill = PriorShape, color = PriorShape))+
    geom_density(aes(y = ..scaled..), size = 0.6, alpha = 0.7)+
    theme_few() +
    scale_x_continuous(breaks = c(0, 1), limits= c(0, 1))+
    scale_y_continuous(breaks = c(0, 1), limits= c(0, 1))+
    ylab("Normalized probability density") +
    xlab("")+
    scale_color_solarized()+
    scale_fill_solarized()+
    facet_grid(PriorShape~., scales = 'free')+
    theme(strip.text.y = element_blank())
```

>>>>

---
title: "Lang-gen model simulations"
output: html_notebook
---

@@

# RSA model

@@

```{r rsaInterpretationModels}
no.utterance.model <- '
var noUtteranceInterpreter = function() {
  Infer({model: function(){
   var state = sample(Beta( flip(mixture) ? priorParams : {a:1,b:100}))
    return { state }
 }, method: "rejection", samples: 20000, burn:5000, verbose: T})}
'

@@

```{r simluationEndorsementModelRuns}
s1.model <- '
var interpreter = cache(function(utterance) {
  Infer({model: function(){
    var state = sample(statePrior)
    var theta = utterance == "generic" ? sample(thetaPrior) : -99
    condition(meaning(utterance, state, theta))
    return state
 }})}, 10000)
